Gaps in stellar streams as a result of globular cluster fly-bys

### 2024-12-11 paper/main.tex
updated conversation about Banik 2017

    We briefly compare our simulated gaps to the literature on Paloamr~5. In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by Fig.~\ref{2023MNRAS.520.5225M}. The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from  Baumgardt's catalog and are thus offset when projected into tail coordinates. Nonetheless, we note that the lengths of the streams are at max 6~kpc on either side of the globular cluster. We apply a threshold of 5.75~kpc on either side to consider the gaps that could occur in currently observed regions of Paloamr~5. This is presented in Fig.~\ref{fig:GapsWithinSight}.
      \begin{figure}
        \centering
        \includegraphics[width=\linewidth]{GapWithinSight.png}
        \caption{The distribution of gaps in absolute distance from the center of mass of Palomar~5's globular cluster. A threshold value of $\sim$~5.75~kpc was selected as distance threshold that corresponds to the current observed length. Distances beyond this are shaded. }
        \label{fig:GapsWithinSight}
      \end{figure}

@@

  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar~5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal~5 streams. From their study, they derived that the impact of globular clusters was marginal in gap creation. It is difficult to give an exact explanation of the reasons for this difference, because the approach followed by  \citet{2019MNRAS.484.2009B} is extremely similar to ours. We would point out, however, that the Galactic potential in which the integration of the orbit of Pal5 and all other galactic clusters was made in their study is the \texttt{MWPotential2014} \citep{2015ApJS..216...29B}. As discussed by \citet{2022ApJ...926..107M}, this potential has a halo of dark matter that is probably too light, because 20\% of the Galactic globular clusters turn out to have unbound orbits in it (note that this is not the case for the potential we have adopted, see e.g. Fig.~\ref{fig:mass_size_plane}). Even for clusters that turn out to be bound in the  \texttt{MWPotential2014} potential, the fact that the dark matter halo is (possibly) too light implies that the orbital volume occupied by the clusters is larger, thus reducing the probability of close impacts with the Pal~5 stream. We tentatively propose this as an explanation of the differences found, but a more systematic comparison between \citet{2019MNRAS.484.2009B}'s results and ours should be made to understand the origin of this difference. \\

>>>>

    We briefly compare our simulated gaps to the literature on Paloamr~5. In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}. The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from  Baumgardt's catalog and are thus offset when projected into tail coordinates. Nonetheless, we note that the lengths of the streams are at max 6~kpc on either side of the globular cluster. We apply a threshold of 5.75~kpc on either side to consider the gaps that could occur in currently observed regions of Paloamr~5. This is presented in Fig.~\ref{fig:GapsWithinSight}.
      \begin{figure}
        \centering
        \includegraphics[width=\linewidth]{GapWithinSight.png}
        \caption{The distribution of gaps in absolute distance from the center of mass of Palomar~5's globular cluster. A threshold value of $\sim$~5.75~kpc was selected as distance threshold that corresponds to the current observed length. Distances beyond this are shaded. }
        \label{fig:GapsWithinSight}
      \end{figure}

@@

  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal5 streams. While their methods are extremely similar to ours, their analysis diverges significantly. Specifically, \citet{2019MNRAS.484.2009B} focused on examining power spectra, analyzing variations in the one-dimensional stream density in stellar counts along the length of the stream.

  Upon comparison with our simulations, we observe that the power spectra from our \texttt{full} simulations that include globular clusters and the \texttt{reference} simulations that do not, are not significantly different. We suggest that this may be due to the signal from only one or two gaps caused by globular clusters not being sufficient to produce notable differences in the overall power spectrum. Additionally, \citet{2022ApJ...926..107M} may not have inspected the profiles for individual gaps, which could explain why these features were not reported in their study. Further investigations would be needed to confirm this interpretation and to thoroughly assess the impact of globular clusters on stream density profiles. \\
